---
title: "In-class Ex 10"
author: "Nathania Yeo"
date: "October 27, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
---

```{r}
pacman::p_load(olsrr, ggstatsplot, sf, tmap, tidyverse, gtsummary, performance, see, sfdep)
```

```{r}
condo_resale <- read_csv("data/aspatial/Condo_resale_2015.csv")
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

```{r}
condo_resale.sf <- st_as_sf(condo_resale, coords = c("LONGITUDE", "LATITUDE"), crs=4326) %>% st_transform(crs = 3414)
```

Building a Hedonic Pricing Model by using Multiple Linear Regression Method

```{r}
condo_mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE    + 
                  PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                  PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET + PROX_KINDERGARTEN + 
                  PROX_MRT  + PROX_PARK + PROX_PRIMARY_SCH + 
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_SUPERMARKET + 
                  PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                data=condo_resale.sf)
summary(condo_mlr)
```

```{r}
ols_condo <- ols_regress(condo_mlr)
```

Variable selection

Stepwise regression is the step-by-step iterative construction of a regression model that invovles the selection of independent variables to be used in the model.

-   ols_step_forward_p

-   ols_step_backward_p

-   ols_step_both_p

```{r}
condo_fw_mlr <- ols_step_forward_p(condo_mlr, p_val=0.05, details=FALSE)
condo_fw_mlr
```

```{r}
plot(condo_fw_mlr)
```

```{r}
condo_bw_mlr <- ols_step_backward_p(condo_mlr, p_val=0.05, details=FALSE)
condo_bw_mlr
```

```{r}
plot(condo_bw_mlr)
```

```{r}
condo_both_mlr <- ols_step_both_p(condo_mlr, p_val=0.05, details=FALSE)
condo_both_mlr
```

```{r}
plot(condo_both_mlr)
```

Model Selection

```{r}
metric <- compare_performance(condo_mlr, condo_fw_mlr$model, condo_bw_mlr$model, condo_both_mlr$model)
```

```{r}
metric$Name <- gsub(".*\\\\([a-zA-Z0-9_]+)\\\\, \\\\model\\\\.*", "\\1", metric$Name)
```

```{r}
plot(metric)
```

```{r}
ggcoefstats(condo_both_mlr$model, sort = "ascending")
```

Check for multicollinearity

```{r}
check_collinearity(condo_both_mlr$model)
```

```{r}
plot(check_collinearity(condo_both_mlr$model)) + theme(axis.text.x = element_text(angle=45, hjust=1))
```

Linearity Assumption Test

```{r}
out <- plot(check_model(condo_both_mlr$model, panel=FALSE))
out[[2]]
```

Normality assumption test

```{r}
plot(check_normality(condo_both_mlr$model))
```

Checking for outliers

```{r}
outliers <- check_outliers(condo_both_mlr$model, method = "cook")
outliers
```

```{r}
plot(check_outliers(condo_both_mlr$model, method="cook"))
```

Visualising spatial non-stationary

```{r}
mlr_output <- as.data.frame(condo_fw_mlr$model$residuals) %>% rename(`FW_MLR_RES` = `condo_fw_mlr$model$residuals`)
```
