---
title: "In-class Ex 4"
author: "Nathania Yeo"
date: "September 9, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
---

```{r}
pacman::p_load(sf, raster, spatstat, sparr, tmap, tidyverse)
```

# Importing Data

## Importing and preparing Study Area Data

```{r}
#| eval: false
kbb <- st_read(dsn = "data/rawdata", layer = "Kepulauan_Bangka_Belitung") 
```

In kbb, there are a few issues that we need to fix. - The geometry type is Polygon Z. If we want to use ppp later on, polygon z will create errors. - Need to drop Z entities - Coordinate system might need to be edited

```{r}
kbb_sf = st_read(dsn = "data/rawdata", layer = "Kepulauan_Bangka_Belitung") %>% 
   st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs=32748)
```

```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```

```{r}
class(kbb_owin)
```

## Importing and Preparing Forest Fire Data
Remove latitude and longitude columns
```{r}
fire_sf <- read_csv("data/rawdata/forestfires.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"),
           crs=4326) %>%
  st_transform(crs=32748)
```
Because ppp object only accept numerical or character as mark. The below code chunk is used to convert data type of acq_date to numeric
```{r}
fire_sf <- fire_sf %>%
  mutate(DayofYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>% 
  mutate(Month_fac = month(acq_date, 
                            label=TRUE, 
                            abbr=FALSE))
```

# Visualising the Fire Points
plot overview first before points
polygons, line, points
```{r}
tm_shape(kbb_sf) + tm_polygons() + tm_shape(fire_sf) + tm_dots()
```

Month_fac is to ensure that the graphs are plotted based on the months in order
free.coords is important, if set to TRUE, zoom level will be be where the data are
drop.units are important as well 
```{r}
tm_shape(kbb_sf) + tm_polygons() + tm_shape(fire_sf) + tm_dots(size=0.1) + tm_facets(by="Month_fac", free.coords=FALSE, drop.units = TRUE)
```

# Computing STKDE by month
## Extract forest fires by month
```{r}
fire_month <- fire_sf %>% 
  select(Month_num)
```

## Creating ppp
```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```
```{r}
summary(fire_month_ppp)
```
```{r}
any(duplicated(fire_month_ppp))
```

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

# Computing Spatio-temporal KDE
```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```
```{r}
#| fig-width: 12
#| fig-height: 10
tims <- c(7, 8, 9, 10, 11, 12)
par(mfcol=c(2, 3))
for(i in tims){
  plot(st_kde, i, override.par=FALSE,
       fix.range=TRUE,
       main=paste("KDE at month", i))
}
```


