---
title: "In-class Ex 11"
author: "Nathania Yeo"
date: "November 4, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
---

# Load Packages

```{r}
pacman::p_load(tidyverse, sf, tmap, httr, performance)
```

# Load Data

```{r}
folder_path <- "data/aspatial"
file_list <- list.files(path=folder_path, pattern="^realis.*\\.csv$", full.names=TRUE)

realis_data <- file_list %>% map_dfr(read_csv)
```

Make sure postal code is 6 digit character

Sale Date is character not numerical (cannot do data manipulation, cannot search by month/weeks) -\> change to a date field

# Wrangling Data

Convert values in Sale Date from character to numerical

Extract only resale condo

```{r}
condo_resale <- realis_data %>% mutate(`Sale Date` = dmy(`Sale Date`)) %>% 
  filter(`Type of Sale` == "Resale" & `Property Type` == "Condominium")
```

# Geocoding

```{r}
postcode <- unique(condo_resale$'Postal Code')
```

```{r}
url <- "https://onemap.gov.sg/api/common/elastic/search"
found <- data.frame()
not_found <- data.frame()

for (postcode in postcode){
  query <- list('searchVal'=postcode, 'returnGeom'='Y',
                'getAddrDetails'='Y', 'pageNum'='1')
  res <- GET(url, query=query)
  if ((content(res)$found)!=0){
    found <- rbind(found, data.frame(content(res))[4:13])
  } else {not_found = data.frame(postcode)
  }
}
```

# Tidying field names

```{r}
found <- found %>% select(c(6:8)) %>% rename(POSTAL='results.POSTAL', 
                                             XCOORD = 'results.X', 
                                             YCOORD = 'results.Y')
```

# Join tables

```{r}
condo_resale_geocoded = left_join(condo_resale, found, by =c("Postal Code"= "POSTAL"))
```

```{r}
condo_resale_sf <- st_as_sf(condo_resale_geocoded, coords=c("XCOORD", "YCOORD"), crs=3414)
```

# Cleaning the spatial data

Checking for overlapping point features

```{r}
overlapping_points <- condo_resale_sf %>% mutate(overalp = lengths(st_equals(., .))>1)
```

Spatial jittering

-   Don't use values such as 0.5 or 0.1 for amount parameter

```{r}
condo_resale_sf <- condo_resale_sf %>% st_jitter(amount=2)
```
